---
title: "Generating Hydrometric Reports with StreamTrackR"
output: 
  github_document:
    html_preview: true
always_allow_html: true
---

# **Overview**
**Version:** 1.0.0  
**Last Edited:** `r Sys.Date()`   
**Author:** Paula Soto    
**Email:** paula.soto@dfo-mpo.gc.ca

This vignette demonstrates how to use the `streamTrackR` project to generate hydrometric condition reports. The workflow utilizes an RMarkdown template and specified parameters to produce an HTML report using Water Survey of Canada (WSC) station data.

### Setup Instructions
1. Clone the project from the repository using your preferred method. Below is one example of how to do it in the R Studio terminal:

```bash
git clone https://github.com/pausoto7/StreamTrackR.git
```
   

2. Install required R packages 

Install `hydroGraphR` from GitHub. `hydroGraphR` is an R package designed to simplify the process of creating hydrographs. 
```{r hydrog, eval = FALSE}
# Install devtools if not already installed
install.packages("devtools")

# Install hydroGraphR
devtools::install_github("pausoto7/hydroGraphR")

```

Install the rest of the required R packages from CRAN:

```R
install.packages(c("ggplot2",  "formattable", "knitr", "kableExtra",
                       "rmarkdown", "lubridate", "dplyr", "tidyr"))
```

3. Open the project in RStudio and open `run_hydro_report.R` to begin. 


# **Getting started**

## Setting Up 

The first step is to download the required libraries and create an isolated environment for your script to run in.

```{r createenv, results = 'hide', warning=FALSE, message=FALSE}

library(tidyhydat)
library(hydroGraphR) 
library(dplyr)

# Create a new environment for rendering 
isolated_env <- new.env()

```


## Define the file name and output directory

Next, define what the name of your file will be. In this example the report will be for conditions in the Yukon. Any information may be included here, but the file must end in ".html". In this example information on the location of the report is included along with the date it is run (today). Other information that could be helpful to include would be whether the report will be output in calendar or water year format. 

The following lines add a "Reports" folder into the project if one doesn't already exist. 

```{r filename, results = 'hide', warning=FALSE, message=FALSE}

file_name <- paste0("Reports/Yukon Condition Report", Sys.Date(), ".html")

# Add a "Reports" folder if it doesn't already exist
if (!dir.exists("Reports")) 
  { dir.create("Reports") }

```
## Rendering the Report

The `rmarkdown::render` function is used to generate the report. The function takes the following parameters:

- **Stations**: A list of WSC station IDs for stations that should be include in the report. Station ID's can be found on the [WSC website](https://wateroffice.ec.gc.ca/search/real_time_e.html). 
- **YOI**: The year of interest for the data.
- **Location**: The location used for the title of the report.
- **WY**: Logical value indicating whether to present hydrograph by water year (Nov-Oct)(`TRUE`) or calendar year (Jan-Dec)(`FALSE`).

```{r rmd, eval = FALSE}
# Render the R Markdown report
rmarkdown::render(
  here::here("hydrometric_report.Rmd"),
  output_file = file_name,
  params = list(
    stations = c(
      "09AG001", "09EA006", "09CA002"
    ),
    YOI = 2025,
    location = "Yukon", # To be used for the title of the report
    WY  = FALSE
  ),
  envir = isolated_env
)
```

## Viewing the Report

Once the rendering process is complete, the report will be saved in the `Reports` directory. Open the HTML file to view the results.

Note that all the figures produced in the report will also be present in the similarly named folder within the same `Reports` directory. 

<br>

# *Example Report Output:*


```{r report title, results='asis', include=TRUE, echo=FALSE}
 cat('\n\n<!-- -->\n\n')

# Create the dynamic title and date
report_title <- sprintf('<h1 style="font-size: 2.8em;">%s Hydrometric Conditions Report</h1>', "Yukon")
system_date <- sprintf('<h2 style="font-size: 1.8em; color: #5b5b5b;">%s</h2>', format(Sys.Date(), '%B %d, %Y'))

# Print the title and date
cat(report_title, "\n")
cat(system_date, "\n")
```

##### This report provides an overview of current hydrologic conditions at selected stations. Level and flow data are sourced from both historical and real-time hydrometric data published by Water Survey of Canada. See below table for terms and definitions used in station tables. 



```{r createDefTable, echo=FALSE, results='asis'}


# Define custom colors
green_bg <- "#bef7a6"  # Light green background
red_bg <- "#f7a6a6"    # Light red background


definitions <- data.frame(
  Term = c("Station Number", "Parameter", "Mean (Today)",
           "Historical Mean (Today)", "Percent of Historical",
           "MAD", "Q25 (Today)", "Q75 (Today)", "72 Hr Change",
           "Date Ranges", "Trajectory"),
  
  Definition = c(
    "WSC station number with a link to the WSC station website.", 
    "Parameter type, either flow or water level.",
   paste(
      "The average value of today's (or most recent date captured) measurements. ",
      "<span style='background:", green_bg, ";'>Green</span> symbology indicates the value is between Q25-Q75. ",
      "<span style='background:", red_bg, ";'>Red</span> indicates the value is out of this range."
    ),    "The average value of the measurements on this date over the historical period.",
    "The percentage of today's value compared to the historical average.",
    "Mean Annual Discharge, a measure of the average amount of discharge through a river or stream over the course of a year.",
    "The 25th percentile of today's data distribution, indicating that 25% of the data points are below this value.",
    "The 75th percentile of today's data distribution, indicating that 75% of the data points are below this value.",
    "The change in the measurement over the past 72 hours.",
    "The years during which the data was collected and included in analysis.",
    "The trend or direction of the measurement over a three day period."
  )
)

# Create the kable table
knitr::kable(definitions, "html", escape = FALSE, col.names = c("Term", "Definition")) %>%
  kableExtra::kable_styling(full_width = TRUE, position = "left", bootstrap_options = "condensed") %>%
  kableExtra::row_spec(0, bold = TRUE, background = "#f0f0f0")

  #cat("<br>") # Adds extra space

```

```{r stations data, include = FALSE,  results='asis'}

# gradient_color -----------------------------------

format_color <- function(x) {
  ifelse(x < 80, formattable::color_tile("red", "#FFFFEE")(x), color_tile("#EEFFEF", "olivedrab")(x))
}

# ---------------------------------------------------


format_mean <- function(x, Q25, Q75) {
  formattable::formatter("span", style = ~ formattable::style(
    background = ifelse(x >= Q25 & x <= Q75, "#bef7a6", "#f7a6a6")
  ))(x)
}


```


```{r example, echo=FALSE, results='asis',  out.width='100%', fig.height = 4, message = FALSE, warning = FALSE}

all_stations_data <- hydroGraphR::dl_hydro("09AG001")

st <- "09AG001"
YOI <- 2024 
WY <- TRUE
source("R/utils.R")

max_date <-as.Date(paste(as.character(YOI-1), "12", "31", sep = "-"))

stats_table_info<- create_stats_table(all_stations_data, YOI = YOI)
  
stats_table_raw <- stats_table_info[[1]]
  
cat('\n\n<!-- -->\n\n')
cat(paste0("### ", hy_stations(st)$STATION_NAME, "\n\n"))
cat('\n\n<!-- -->\n\n')

stats_table <- stats_table_raw %>% select(-station_name)
  

colnames(stats_table) <- c("Station Number", "Parameter", "Mean (Today)",
                             "Historical Mean (Today)", "Percent of Historical", "MAD",
                             "Q25 (Today)", "Q75 (Today)",  "72 Hr Change", "Date Ranges", "Trajectory")
  
# Add formatting to Formatted Mean column which will eventually get changed to Historical Mean
  stats_table <- stats_table %>%
    rowwise() %>%
    mutate(`Formatted Mean (Today)` = list(format_mean(`Mean (Today)`,
                                                     `Q25 (Today)`,
                                                     `Q75 (Today)`)))

# Unnest the formatted column for correct display
  stats_table <- stats_table %>% 
    ungroup() %>%
    mutate(`Formatted Mean (Today)` = sapply(`Formatted Mean (Today)`, as.character))


#  save formattable table to be printed
  formattable_table <- formattable::formattable(stats_table %>% 
                                  select(-`Mean (Today)`, `Q25 (Today)`, `Q75 (Today)`) %>%
                                  rename(`Mean (Today)` = `Formatted Mean (Today)`) %>%
                                   select("Station Number", "Parameter", "Mean (Today)", "Historical Mean (Today)",
                                          "Percent of Historical", "MAD", "Q25 (Today)", "Q75 (Today)",  "72 Hr Change",  "Date Ranges", "Trajectory"))


# Convert to HTML and style the table
  html_table <- knitr::kable(formattable_table,
                           "html",
                           escape = FALSE, 
                           align = "c") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  kableExtra::footnote(general_title = "", 
                       general = paste("Note:\n", stats_table_info[2], "\n   ", stats_table_info[3])) %>%
  kableExtra::row_spec(row = 0, color = "white", background = "#0097A9") 

  cat(knitr::knit_print(html_table))
  

  cat("<br>") # Adds extra space


  # gets stats for flow and WL separately 
  st_df_past <- create_hydro_stats_historical(all_stations_data, date_minimum = NA, date_maximum = max_date)
  sf_df_currentYr <- create_hydro_stats_singleYr(all_stations_data, YOI, WY = WY)

  
  # SET UP HYDROGRAPH ------------------------------------------------

    invisible(hydroGraphR::create_hydrograph_separate(all_hydro_sites_historical = st_df_past,
                                              all_hydro_sites_1yr = sf_df_currentYr, 
                                              parameter = "flow",
                                              output_type = "print", 
                                              WY = WY))

    
  
```
  

# Conclusion

This vignette has demonstrated how to use the `streamTrackR` project to generate a custom hydrometric report. Modify the parameters as needed to customize the stations, year of interest, and location for your specific requirements.

<br>


### To modify this project
If you'd like to modify this project, please **fork** it and make changes in your own version.
If you have improvements, feel free to submit a **pull request**!

<br>

