---
title: "Generating Hydrometric Reports with StreamTrackR"
output: 
  github_document:
    html_preview: true
always_allow_html: true
---

# **Overview**
**Version:** 1.0.0  
**Last Edited:** `r Sys.Date()`   
**Author:** Paula Soto    
**Email:** paula.soto@dfo-mpo.gc.ca

This project provides a summary of current hydrologic conditions at selected stations using both historical records and real-time hydrometric data from the Water Survey of Canada. It includes key statistics for each station and visualizes water level or discharge trends within a historical context.

The report is designed to facilitate quick comparisons between stations while incorporating real-time data and long-term trends.

This vignette outlines the process for generating hydrometric condition reports. Using an RMarkdown template and user-defined parameters, the workflow produces an HTML report based on WSC station data.

### Project Download & Setup Instructions

1. Clone the project from the GitHub repository using your preferred method. 

- One way to do this is to follow the steps found [here](https://docs.github.com/en/repositories/creating-and-managing-repositories/cloning-a-repository). Repository path is [here](https://github.com/pausoto7/StreamTrackR.git).
  


2. Install required R packages

Install `hydroGraphR` from GitHub. [hydroGraphR](https://github.com/pausoto7/hydroGraphR.git) is an R package designed to simplify the process of creating hydrographs. 
```{r hydrog, eval = FALSE}
# Install devtools if not already installed
install.packages("devtools")
library(devtools)

# Install hydroGraphR
install_github("pausoto7/hydroGraphR")

```

Install the rest of the required R packages from CRAN:

```R
install.packages(c("ggplot2",  "formattable", "knitr", "kableExtra", flextable,
                       "rmarkdown", "lubridate", "dplyr", "tidyr", "stringi", "here"))
```

3. Open the project in RStudio and open `run_hydro_report.R`. The file can be found in the R project within the *R* folder. 


# **Getting Started**

## Setting Up 

Once you've opened `run_hydro_report.R`, the first step is to download the required libraries and create an isolated environment for your script to run in. The rest of the libraries are run within the rmarkdown script which is why they're not loaded below. 

```{r createenv, results = 'hide', warning=FALSE, message=FALSE}


library(tidyhydat)
library(tidyverse)

#New environment for rmarkdown:render()
isolated_env <- new.env()

# if Reports folder doesn't already exist, then this creates one
if (!dir.exists("Reports")){
  dir.create("Reports")
}


```


## Define Report Parameters

Before generating the report, key parameters must be defined to specify the location, stations, time period, and formatting preferences for the report. These parameters ensure that the report and report name are both descriptive and relevant to the analysis being conducted.

- **location_name**: this variable specifies the geographic area for the report. This name will be used in both the report title and file name.
- **file_name**: This determines the  reports file name. It is constructed dynamically using `sprintf()`, incorporating the location name, the date the report is run (`Sys.Date()`), and any other specified info. The report should be saved in the "Reports/" directory to keep outputs organized.
- **station_IDs**: A list of WSC station IDs for stations that should be included in the report. Station IDs can be found on the [WSC website](https://wateroffice.ec.gc.ca/search/real_time_e.html). 
- **YOI**: The year of interest for the data.
- **WY**: Logical value indicating whether to present hydrograph by water year (Nov-Oct)(`TRUE`) or calendar year (Jan-Dec)(`FALSE`). For more information on water years see [here](https://www.ausableriver.org/blog/what-water-year#:~:text=Hydrologists%20measure%20time%20differently%20than,known%20as%20a%20water%20year.).



```{r report, results = 'hide', warning=FALSE, message=FALSE}

# LOCATION NAME
location_name <- "Yukon"

# FILE NAME
file_name <- sprintf("Reports/%s Condition Report %s.html", location_name, Sys.Date())

# STATION NAMES
station_IDs <- c("09AG001", "09EA006", "09CA002", "09AE006", "09BC001", "09FD003", "09DD003",
"09DD004", "09AC001",  "09CD001", "09AH001", "09EB001", "09AB001")

# YOI
YOI <- 2025

#WY
WY_type <- TRUE


```

## Rendering the Report

The `rmarkdown::render` function is used to generate the report with the parameters specified above. 



```{r rmd, eval = FALSE}
# Render the R Markdown report

rmarkdown::render(here::here("hydrometric_report.Rmd"),
  output_file = file_name,
  params = list(stations =  stations_to_run , 
                YOI = YOI, 
                location = location_name,  # To be used for title of report
                WY = WY_type),
  envir = isolated_env)

```

## Viewing the Report

Once the rendering process is complete, the report will be saved in the `Reports` directory. Open the HTML file to view the results.

Note that all the figures produced in the report will also be present in the corresponding folder within the same `Reports` directory. 

<br>

# *Example Report Output:*


```{r report title, results='asis', include=TRUE, echo=FALSE}
 cat('\n\n<!-- -->\n\n')

# Create the dynamic title and date
report_title <- sprintf('<h1 style="font-size: 2.8em;">%s Hydrometric Conditions Report</h1>', "Yukon")
system_date <- sprintf('<h2 style="font-size: 1.8em; color: #5b5b5b;">%s</h2>', format(Sys.Date(), '%B %d, %Y'))

# Print the title and date
cat(report_title, "\n")
cat(system_date, "\n")
```

##### This report provides an overview of current hydrologic conditions at selected stations. Level and flow data are sourced from both historical and real-time hydrometric data published by Water Survey of Canada. See below table for terms and definitions used in station tables. 



```{r createDefTable, echo=FALSE, results='asis'}


# Define custom colors
green_bg <- "#bef7a6"  # Light green background
red_bg <- "#f7a6a6"    # Light red background


definitions <- data.frame(
  Term = c("Station Number", "Parameter", "Mean (Today)",
           "Historical Mean (Today)", "Percent of Historical",
           "MAD", "Q25 (Today)", "Q75 (Today)", "72 Hr Change",
           "Date Ranges", "Trajectory"),
  
  Definition = c(
    "WSC station number with a link to the WSC station website.", 
    "Parameter type, either flow or water level.",
   paste(
      "The average value of today's (or most recent date captured) measurements. ",
      "<span style='background:", green_bg, ";'>Green</span> symbology indicates the value is between Q25-Q75. ",
      "<span style='background:", red_bg, ";'>Red</span> indicates the value is out of this range."
    ),    "The average value of the measurements on this date over the historical period.",
    "The percentage of today's value compared to the historical average.",
    "Mean Annual Discharge, a measure of the average amount of discharge through a river or stream over the course of a year.",
    "The 25th percentile of today's data distribution, indicating that 25% of the data points are below this value.",
    "The 75th percentile of today's data distribution, indicating that 75% of the data points are below this value.",
    "The change in the measurement over the past 72 hours.",
    "The years during which the data was collected and included in analysis.",
    "The trend or direction of the measurement over a three day period."
  )
)

# Create the kable table
knitr::kable(definitions, "html", escape = FALSE, col.names = c("Term", "Definition")) %>%
  kableExtra::kable_styling(full_width = TRUE, position = "left", bootstrap_options = "condensed") %>%
  kableExtra::row_spec(0, bold = TRUE, background = "#f0f0f0")

  #cat("<br>") # Adds extra space

```

```{r stations data, include = FALSE,  results='asis'}

# gradient_color -----------------------------------

format_color <- function(x) {
  ifelse(x < 80, formattable::color_tile("red", "#FFFFEE")(x), color_tile("#EEFFEF", "olivedrab")(x))
}

# ---------------------------------------------------


format_mean <- function(x, Q25, Q75) {
  formattable::formatter("span", style = ~ formattable::style(
    background = ifelse(x >= Q25 & x <= Q75, "#bef7a6", "#f7a6a6")
  ))(x)
}


```


```{r example, echo=FALSE, results='asis',  out.width='100%', fig.height = 4, message = FALSE, warning = FALSE}

all_stations_data <- hydroGraphR::dl_hydro("09AG001")

st <- "09AG001"
YOI <- 2024 
WY <- TRUE
source("R/utils.R")

max_date <-as.Date(paste(as.character(YOI-1), "12", "31", sep = "-"))

stats_table_info<- create_stats_table(all_stations_data, YOI = YOI)
  
stats_table_raw <- stats_table_info[[1]]
  
cat('\n\n<!-- -->\n\n')
cat(paste0("### ", hy_stations(st)$STATION_NAME, "\n\n"))
cat('\n\n<!-- -->\n\n')

stats_table <- stats_table_raw %>% select(-station_name)
  

colnames(stats_table) <- c("Station Number", "Parameter", "Mean (Today)",
                             "Historical Mean (Today)", "Percent of Historical", "MAD",
                             "Q25 (Today)", "Q75 (Today)",  "72 Hr Change", "Date Ranges", "Trajectory")
  
# Add formatting to Formatted Mean column which will eventually get changed to Historical Mean
  stats_table <- stats_table %>%
    rowwise() %>%
    mutate(`Formatted Mean (Today)` = list(format_mean(`Mean (Today)`,
                                                     `Q25 (Today)`,
                                                     `Q75 (Today)`)))

# Unnest the formatted column for correct display
  stats_table <- stats_table %>% 
    ungroup() %>%
    mutate(`Formatted Mean (Today)` = sapply(`Formatted Mean (Today)`, as.character))


#  save formattable table to be printed
  formattable_table <- formattable::formattable(stats_table %>% 
                                  select(-`Mean (Today)`, `Q25 (Today)`, `Q75 (Today)`) %>%
                                  rename(`Mean (Today)` = `Formatted Mean (Today)`) %>%
                                   select("Station Number", "Parameter", "Mean (Today)", "Historical Mean (Today)",
                                          "Percent of Historical", "MAD", "Q25 (Today)", "Q75 (Today)",  "72 Hr Change",  "Date Ranges", "Trajectory"))


# Convert to HTML and style the table
  html_table <- knitr::kable(formattable_table,
                           "html",
                           escape = FALSE, 
                           align = "c") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  kableExtra::footnote(general_title = "", 
                       general = paste("Note:\n", stats_table_info[2], "\n   ", stats_table_info[3])) %>%
  kableExtra::row_spec(row = 0, color = "white", background = "#0097A9") 

  cat(knitr::knit_print(html_table))
  

  cat("<br>") # Adds extra space


  # gets stats for flow and WL separately 
  st_df_past <- hydroGraphR::create_hydro_stats_historical(all_stations_data, date_minimum = NA, date_maximum = max_date)
  sf_df_currentYr <- hydroGraphR::create_hydro_stats_singleYr(all_stations_data, YOI, WY = WY)

  
  # SET UP HYDROGRAPH ------------------------------------------------

    invisible(hydroGraphR::create_hydrograph_separate(all_hydro_sites_historical = st_df_past,
                                              all_hydro_sites_1yr = sf_df_currentYr, 
                                              parameter = "flow",
                                              output_type = "print", 
                                              WY = WY))

    
  
```
  

# Conclusion  

This vignette has demonstrated how to use the `streamTrackR` package to generate a **custom hydrometric report**. By adjusting key parameters—such as **station selection, year of interest, and location**—users can tailor reports to their specific needs.  

Beyond the basic functionality covered here, `streamTrackR` provides flexibility for further customization, including **modifying visualizations, adjusting data formats, and integrating additional hydrometric datasets**.  

To explore more features, experiment with different parameter settings to generate reports that best suit your hydrological analysis needs.  
<br>


### To modify this project
If you'd like to modify this project, please **fork** it and make changes in your own version.
If you have improvements, feel free to submit a **pull request**! 

<br>

